Speculation in foreign exchange: noise or information

### 2014-11-14 Knitr/Paper.tex
Completed the review of eventdoc.  Now needs to be re-read and the conclusions need to be drawn towards the front.  Some similification of the literature.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\usepackage[dcucite]{harvard}
\usepackage{tikz}
\usetikzlibrary{shapes, shadows, arrows}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\bibliographystyle{plainnat}
\usepackage{listings}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
\begin{document}
%\renewcommand[\harvardurl]{URL: \url}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.  We are very grateful for the comments and suggestions made.  In particular, the EACES conference in Budapest between September 5th and 6th 2014, Andros Steinhoffer and.....}} 
\date{\today}
\maketitle
\begin{abstract}
An event study of extreme speculation is used to test the informational content of speculation.  Speculative extremes are ranked by weight of activity and intensity of sentiment.  In either case, the extreme says little about the future, suggesting that speculation is more than just noise and that there is some fundamental informational content. A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.  This is suplemented with regulatory data on the futures position of speculative traders in the fugtures market.  
\end{abstract}

@@

One of the many things that divided John Maynard Keynes and Milton Friedman was their opinion of speculation.   Keynes main discussion of speculation was in Chapter 13 of \emph{The General Theory} where he identified speuclation as a myopic, sentiment-driven activity that was driven by the desire to ``beat the gun'' or  ``outwit the crowd'' \citep[p. 101]{Keynes1936}.  Friedman, however, say speculation as a stabilising force that would ensure that markets swiftly found equilibrium. In arguing in favour of flexible exchange rates, Friedman asserts that speculators have a stabilsing influence as their success and domination will depend on correctly buying when prices are below fundamental value and selling when above the correct rate.

In Keynes' view, speculative activity will tend to drive prices away from fundamental value as price changes and sentiment feed off each other.  At its most extreme, speculation is likely to cause a series of bubbles.  However, the alternative view regards speculation as an activity that is driven by fundamental knowledge and information and acts to limit overshooting and divergence from fundamental value, thereby reducing volatility.  This second view says that it is the absence of speculators and the liquidity that they provide that encourages excess price movement and volatility.

The aim of this paper is to use the foreign exchange market to assess the relationship between intensity or weight of speculation and the movement of prices.  If speculators are uninformed, their most extreme sentiment and their greatest weight in market activity should tend to cause overshooting and presage bubble-bursting, reversals and a crash; if speculators are informed, activity will drive prices towards fundamental value and equilibrium where volatility and crash risk should be reduced.

From "How noise trading..." in the other folder. 
O'Hara and Saar suggest that there is a difference between noise traders and liquidity traders.  Noise traders exert positive effects on liquidity but weaken price reversals. A tax on trading will reduce noise traders but will also reduce the influence of fundamental traders.

@@

\subsection{Event studies}
The event studies method ???? of how the extreme speculative sentiment and extreme speculative positions in the DeLong model could be used to analyse future returns.  Fuller explanation is provided here. Event studies are used to find the effect of a particular event on the performance of an asset price or security.  It is often suggested that James Dolley was the first to employ the event study when looking at the effect of a stock split on the price of securities.\footnote{A stock split is when the number of shares are doubled and the value of each share is cut in half.  This is usually done to prevent the price or cost of each share rising too far beyond the means of the small investor.}   Dolley took a sample of 174 stocks splits for all common stocks listed on the New York Stock Exchange during the period 1921 to 1931, and using what he called a case study, examined the process of stock splitting.  Price information was taken from \emph{Poor's Manual} and \emph{Moody's Manual} as well as the \emph{New York Times}; a survey was sent to all the companies that were involved and (remarkably) 63 of the 111 questionnaires were returned completed.  The analysis of the price change was a very small part of the study.  Dolly found that of the 95 stock splits that he had been able to compile a time series for, 57 subsequently rose and 26 fell \citep{Dolly1933}.

However, Dolly only looked at the market reaction the day after the stock split.  This was some time after the stock holder meeting that would have ratified the decision of the management. Myer and Bakay argued that, if stocks tend to rise after a stock split, it would be best to buy the stocks when the announcement is made and before the split takes place and that, therefore, the price adjustment is likely to be spread over a number of days.  They decided to measure the price action from a point eight weeks before the formal acceptance of the decision by the stock-holders. They measured the price change between the base date and the meeting and tried to account for other influences on the price by recording the ratio of the stock price to the Standard and Poor Industry average, thereby providing the first attempt to separate the specific returns from those of the market.  An additional measure of the price ratio was also taken eight weeks after the formal meeting.  For the splits that occurred between 1945 and 1946 they found that 62 of the 70 stocks exhibited a positive price performance between the base date and the meeting, during the subsequent time there was some reversal with only 55 of the stocks still showing a net gain between the base and the closure point eight weeks after the announcement meeting \citep{MyersBakay}.

Fama, Fisher and Roll use the event study method to conduct a much more systematic assessment of the effect of stock splits on security returns \citep{FamaFisherJensenRoll}.    Fama later said that the method was secondary to his desire to show off the new dataset that he had helped to construct.  This was one of the first uses of the Center for Research in Security Prices (CRSP) data.  Nonetheless, the Fama study is considered to be the benchmark of Event Studies and it is the framework that is largely followed in the current work.

The event is identified and the time at which the event occurs is marked as time zero (or $t_0$).  Returns to the reference security in a window before and after the event are recorded and compared to the normal or expected returns.  Therefore, 
\begin{equation}
AR_{i \tau} = R_{i \tau} - E[R_{i \tau} | X_{\tau}]
\end{equation}
Where $AR_{i \tau}$ are the abnormal returns for security i at time $\tau$, $R_{i \tau}$ are the actual returns to security i at time $\tau$,  $E[R_{i \tau}]$ are the expected returns for security i at time $\tau$ conditional on the information $X_{\tau}$ and $\tau$ is the time relative to the event.  Estimating the normal or expected returns is one of the key challenges of an event  study.  It is usually tackled in one of two ways:  the security can be assumed to take a mean return or, as in the case of the paper by Fama and the others, a market model can be used.  The original market model just takes a constant relationship between security returns and the market estimated by a linear regression.  Subsequent models have added additional factors.

The mean return is calculated as 
\begin{equation}
\mu_{R, i} = \frac{1}{n}\sum_{j = 1}^{j = n} R_{i j}
\end{equation}

where $R_{i j}$ is the return to security i at time j and n is the number of total number of observations to that 
\begin{equation}
E[R_{i \tau} | X_{\tau}] = \mu_{R, i} + \varepsilon_{i, \tau}
\end{equation} 
with $\varepsilon_{i, \tau}$ being a random variable that is assumed to be normally distributed with a zero mean and a standard deviation of $\sigma_{R, i}$. 
The simple market model is estimated by 
\begin{equation}
R_{i, j} = \alpha + \beta_i R_{m, j} + \varepsilon
\end{equation}

Where the $\beta_i$ is the relationship between the market return $R_{m, j}$ and the return on the security of interest $R_{i, j}$ and $\varepsilon$ is a random variable that represents all the idiosyncratic factors that affect the returns to the security.

The lack of any clear benchmark for currency returns and the fact that theoretically exchange rate returns should be on average equal to the forward rate, have made the case for a mean model.  The seminal event studies have used equities and a market model.  The expected return is zero and any returns that differ from that are identified as abnormal. Theoretically, the forward rate should be used but given the high frequency of the data, the interest rate differential is ignored.   The evidence about returns in Tables \ref{tabref:DS1} and \ref{tabref:DS3} suggests that this is not likely to be a major problem.

@@

Again, the most prominent observation is positive association between sentiment and returns.  Intuitively, this is not surprising.  However, it does suggest that foreign exchange prices are not always driven by a random walk. There is clear evidence here of serial correlation in the returns. As sentiment rises (as measured by the increase in the risk-reversal) there is a price appreciation. However, the adjustment is not instantaneous.  This is an example of the sort of trending behaviour that is common in the literature (as reported in Section \ref{secref:empiric}).  It is likely to be at least partially explained by the sort of gradual adjustment to information that was identified as being a consequence of the conservatism heuristic in Section \ref{sec:behavioural} on behavioural finance.

@@

The second heuristic that was discussed in Section \ref{sec:behavioural} was representitiveness.  This was the tendency of decision-makers to focus on the immediate and to forget base probabilities.  It would suggest that investors would think too much about the new information that has driven prices towards a new equilibrium in Figures \ref{fig:ES1} and \ref{fig:ES2} not sufficient to the broad background where other currencies may be more or less attractive.  In practical terms, it suggests that there will be overshooting and that this overshooting will cause reversals when it becomes extreme.  There is little evidence here of any reversal.  After the extremes are reached, there seems more likely to be a period of consolidation or a fluctuation around the new equilibrium.  This is what would be expected if speculation were part of the process of getting information into the price. Once this has taken place, there is consolidation.

@@

The R code for all the event window figures is available in Appendix \ref{AppendixB} Section \ref{Asecref:event}.  This merely takes the data that was used to construct Tables \ref{tabref:RR1} and \ref{tabref:SP1} and puts it in graphical form.  The event windows are stretched to give a clear view of the evolution of exchange rates  around the event.  Confidence intervals are added.  Once event objects FX.exc have been created, parameter for exchange rate, extreme measure, quantile, window length and the high or low indicator can be changed for alternative assessments of speculative extremes. Adjustments to these parameters show that these results are not a function of the choices presented in the main text.  They are robust to alternative specifications.

@@

\section{Speculation and foreign exchange returns}
This chapter has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.  The measurement of speculation takes two forms:  a measurement of the intensity of belief with the risk-reversal skew, and a measurement of speculative positions.  Neither of these is ideal, in the sense of being a random sample from a population.  However, the option prices reveal the intensity of feeling through the weight of money in a speculative part (options) of a speculative market (foreign exchange).  The commitment of traders report records only speculators in the US futures market.  However, this is an important market, there is no reason to believe that US or speculators in the futures market are systematically different from those in the rest of the world and presumably if there were a major difference there would be an arbitrage opportunity between the two markets.

@@

\bibliography{myrefs}

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\usepackage[dcucite]{harvard}
\usepackage{tikz}
\usetikzlibrary{shapes, shadows, arrows}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\bibliographystyle{plainnat}
\usepackage{listings}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
\begin{document}
%\renewcommand[\harvardurl]{URL: \url}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.  We are very grateful for the comments and suggestions made.  In particular, the EACES conference in Budapest between September 5th and 6th 2014, Andros Steinhoffer and.....}} 
\date{\today}
\maketitle
\begin{abstract}
An event study of extreme speculation is used to assess the Keynesian and Friedman views of speculation.  This is partly a test of the informational content of speculation.  Speculative extremes are ranked by weight of activity and intensity of sentiment.  In either case, the extreme says little about the future, suggesting that speculation is more than just noise and that there is some informational content that is being passed through to the price. A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.  This is suplemented with regulatory data on the futures position of speculative traders.  
\end{abstract}

@@

There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of the spectrum is the view of Keynes from Chapter 13 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citep[p. 101]{Keynes1936}.  However, towards the other extreme, Milton Friedman argued that speculation was a stabilising force that would ensure that markets swiftly found equilibrium. In arguing in favour of flexible exchange rates, Friedman asserted that successful speculators would be those that managed to correctly buy when prices  were below fundamental value and sell above that rate \citep{FriedmanPositive}.

This is a contrast between an idea of speculative activity as a nuisance that, at its most extreme, will likely to cause bubbles and reversal and will discourage fundamental long-term investors from participating in the market.  It is noise.   that does not emerge randomly, but feeds off itsef as a feedback of sentiment  encourages  trends and serial correlation in asset prices returns.  The alternative view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.

The aim of this paper is to understand mroe about the nature of speculation in the foreign exchange market by assess the relationship between the intensity or the weight of speculation and the movement of prices.  If speculators are uninformed and their activity can be regarded as noise, their most extreme sentiment and the time when speculators have the greatest weight in market activity should be those when price reversals are most likely to take place.  An event study is used to assess the relationship between speculative extremes and foreign exchange returns.  A unique set of option prices are used to measure the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market.

\section{Literature}
From "How noise trading..." in the other folder. 
\citet{bloomfield2009noise} suggests that there is a difference between noise traders and liquidity traders.  Noise traders exert positive effects on liquidity but weaken price reversals. A tax on trading will reduce noise traders but will also reduce the influence of fundamental traders. Their experiment distinguishes between informed traders, noise traders and liquidity traders.  \emph{SOES bandits} are those taking advantage of the SOES system to trade quickly and take advantage of small price movements. There is a lot of focus on the effect that taxation may have on the different categories of trader. There is an experiment that allows much more to be held constant. There are three views of noise traders that are tested here:  noise traders are skill experts taking advantage of information in the order book or price movements;  rationally providing liquidity by acting as market-maker; acting irrationally as noise traders. Noise traders in this experiement appear to act as irrational contrarian traders. Noise traders refuce the efficiency of the market but increase liquidity.

Noise is related to rational expectations.  A mean zero, normally-distributed random variable influences supply citet{Muth1961Rational}.  Why do noise traders exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders. The effect of noise traders on liquidity is unclear. They may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existant) informational advantages. Noise traders may increase (if they have some priviledged information, they add to the price pressure) or reduce efficiecy (if they increase liquiity and reduce the ratio of informed to uninformed traders).

Are day traders profitable? Are these the face of noise traders?  How does this show up in teh foreign exchange market?

\citet{da2010sum} Seek to measure invetor sentiment and assess how it affects securities prices. They create a index of investor sentiment from web searches and use this to predict return reversals.  The resutls show that sentiment can predict market turns. This is a contrast to traditional ways of measuring sentiment:  proxy for sentiment with market-based measures like volume, closed-end fund discount, IPO first-day returns, IPO volume, option implied volatities and mutual fund flows. Baker and Wurgaler (2007) have a comprehensive survey. This is more than just investors sentiment. There is an issue with the measurment of an input (sentiment) with an output (market price). Use of surveys like the University of Michigan consumer sentiment survey etc.  There is a negative contemporaneous correlation between FEARS and stock market returns. However, there is a reversal in the relationship after two days.  This is consistent with ``sentiment induced temporary mispricing'' \citet[p. 4]{da2010sum}.  There is a positive correlation between the FEARS index and bond prices.  However, thsi also reverses in following days. There is a relatioknship betwen the FEARS index and daily market volatility (realised volatlity and the VIX index). There is also evidence of inflows and outflows from mutual funds.  This may help to explain the contrast between evidence from the equity market and the bond market.

@@

\subsection{Event studies}
Event studies are used to find the influence of a particular event on the performance of an asset price or security.  James Dolley is credited with the first use of the method when looking at the effect of a stock split on the price of securities.\footnote{A stock split is when the number of shares are doubled and the value of each share is cut in half.  This is usually done to prevent the price or cost of each share rising too far beyond the means of the small investor.} \citep{Dolly1933}. However, Dolly only looked at the market reaction the day after the stock split.  This was some time after the stock holder meeting that would have ratified the decision of the management. Myer and Bakay argued that, if stocks tend to rise after a stock split, it would be best to buy the stocks when the announcement is made and before the split takes place and that, therefore, the price adjustment is likely to be spread over a number of days.  They decided to measure the price action from a point eight weeks before the formal acceptance of the decision by the stock-holders. Myer and Barkay measured the price change between the base date and the meeting and tried to account for other influences on the price by recording the ratio of the stock price to the Standard and Poor Industry average, thereby providing the first attempt to separate the specific returns from those of the market \citep{MyersBakay}.

Fama, Fisher and Roll use the event study method to conduct a much more systematic assessment of the effect of stock splits on security returns \citep{FamaFisherJensenRoll}. the study is considered to be the benchmark of Event Studies and it is the framework that is largely followed in the current work.

The event is identified and the time at which the event occurs is marked as time zero (or $t_0$).  Returns to the reference security in a window before and after the event are recorded and compared to the normal or expected returns.  Therefore, 
\begin{equation}
AR_{i \tau} = R_{i \tau} - E[R_{i \tau} | X_{\tau}]
\end{equation}
Where $AR_{i \tau}$ are the abnormal returns for security i at time $\tau$, $R_{i \tau}$ are the actual returns to security i at time $\tau$,  $E[R_{i \tau}]$ are the expected returns for security i at time $\tau$ conditional on the information $X_{\tau}$ and $\tau$ is the time relative to the event.  Estimating the normal or expected returns is one of the key challenges of an event  study.  It is usually tackled in one of two ways:  the security can be assumed to take a mean return or, as in the case of the paper by Fama and the others, a market model can be used.  The original market model just takes a constant relationship between security returns and the market estimated by a linear regression.  Subsequent models have added additional factors.   
The lack of any clear benchmark for currency returns and the fact that theoretically exchange rate returns should be on average equal to the forward rate, have made the case for a mean model.  The seminal event studies have used equities and a market model.  The expected return is zero and any returns that differ from that are identified as abnormal. Theoretically, the forward rate should be used but given the high frequency of the data, the interest rate differential is ignored.   The evidence about returns in Tables \ref{tabref:DS1} and \ref{tabref:DS3} suggests that this is not likely to be a major problem.

Is more needed here?

@@

@@

@@

The event windows are stretched to give a clear view of the evolution of exchange rates  around the event and to ensure that results are not a function of the size of the event window.  Confidence intervals are added.  Once event objects FX.exc have been created, parameter for exchange rate, extreme measure, quantile, window length and the high or low indicator can be changed for alternative assessments of speculative extremes. Adjustments to these parameters show that these results are not a function of the choices presented in the main text.  They are robust to alternative specifications.

@@

\section{Speculation and foreign exchange returns}
This paper has sought to measure speculation and to use this measure to discover more about the relationship between speculation, information and asset prices.  The measurement of speculation takes two forms:  a measurement of the intensity of belief with the risk-reversal skew, and a measurement of speculative positions.  Neither of these is ideal, in the sense of being a random sample from a population.  However, the option prices reveal the intensity of feeling through the weight of money in a speculative part (options) of a speculative market (foreign exchange).  The commitment of traders report records only speculators in the US futures market.  However, this is an important market, there is no reason to believe that US or speculators in the futures market are systematically different from those in the rest of the world and presumably if there were a major difference there would be an arbitrage opportunity between the two markets.

@@

\bibliography{../../myrefs}
